{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} \begin{tabular}{c} {\em or: How I Learned to Stop Worrying and Love the Exponential Explosion} \end{tabular}

### 2018-10-08 thesis/overview.tex
new quote for bgnd
syntax hilight code figures in bgnd

\inspirationalquote{We pray to our ancestors. We've met all of them. None of them were better than anyone else.}
{Mina, Iconoclasts}

@@

\begin{figure}[t]
	\begin{tabular}{cc}
		\begin{tabular}{p{0.45\textwidth}p{0.5\textwidth}}
			{\footnotesize
			\begin{tabular}{l}
				\texttt{int x;} \\
				\texttt{void count() \{} \\
				\texttt{~~~~for (int i = 0; i < 1000; i++)} \\
				\texttt{~~~~~~~~x++;} \\
				\texttt{\}} \\
				\texttt{void main() \{} \\
				\texttt{~~~~tid1 = thr\_create(count);} \\
				\texttt{~~~~tid2 = thr\_create(count);} \\
				\texttt{~~~~thr\_join(tid1);} \\
				\texttt{~~~~thr\_join(tid2);} \\
				\texttt{~~~~printf("\%d\textbackslash{}n", x);} \\
				\texttt{\}} \\
			\end{tabular}
			}
			&
			{\footnotesize \begin{tabular}{ll}
				{\bf \normalsize Thread 1} & {\bf \normalsize Thread 2} \\
				\hline
				\texttt{load tmp <- x;} & \\
				& \texttt{load tmp <- x;} \\
				& \texttt{add tmp <- 1;} \\
				& \texttt{store x <- tmp;} \\
				\texttt{add tmp <- 1;} & \\
				\texttt{store x <- tmp;} & \\
			\end{tabular}
			}
			\\
			\\
			(a) Source listing for a multithreaded program which might count to 2000.
			&
			(b) Example interleaving of the compiled assembly for (a),
			in which 2 concurrent iterations of the loop yield 1 net increment of {\tt x}.
		\end{tabular}
	\end{tabular}
	\caption{Example concurrent program in which simultaneous accesses to shared state may interleave to produce unexpected results.}
	\label{fig:concurrency-bug}
\end{figure}

@@

\begin{figure}[t]
	\begin{tabular}{p{0.45\textwidth}p{0.5\textwidth}}
		{\footnotesize
		\begin{tabular}{l}
			\texttt{typedef struct mutex \{} \\
			\texttt{~~~~volatile int held;} \\
			\texttt{~~~~int owner;} \\
			\texttt{\} mutex\_t;} \\
			\texttt{void mutex\_lock(mutex\_t *mp) \{}\\
			\texttt{~~~~while (xchg(mp->held, 1))} \\
			\texttt{~~~~~~~~yield(mp->owner);} \\
			\texttt{~~~~mp->owner = gettid();} \\
			\texttt{\}} \\
			\texttt{void mutex\_unlock(mutex\_t *mp) \{}\\
			\texttt{~~~~mp->owner = -1;} \\
			\texttt{~~~~mp->held = 0;} \\
			\texttt{\}} \\
		\end{tabular}
		}
		&
		{\footnotesize
		\begin{tabular}{l}
			\texttt{int x;} \\
			\texttt{mutex\_t m;} \\
			\texttt{void count() \{} \\
			\texttt{~~~~for (int i = 0; i < 1000; i++) \{} \\
			\texttt{~~~~~~~~mutex\_lock(\&m);} \\
			\texttt{~~~~~~~~x++;} \\
			\texttt{~~~~~~~~mutex\_unlock(\&m);} \\
			\texttt{~~~~\}} \\
			\texttt{\}} \\
		\end{tabular}
		}
		\\
		(a) A simple mutual exclusion lock built using the {\tt xchg} instruction. %(accessed using a GCC compiler intrinsic).
		&
		(b) The {\tt count} function from Figure~\ref{fig:concurrency-bug}, adjusted to use a mutex to ensure each increment of {\tt x} is uninterruptible.
	\end{tabular}
	\caption{Using a locking primitive to protect accesses to shared state.}
	\label{fig:mutex}
\end{figure}

@@

\begin{figure}[h]
	\begin{center}
		\begin{tabular}{l}
		\texttt{\ctype{void} \call{count}() \{} \\
		\texttt{~~~~\flow{for} (\ctype{int} i = \const{0}; i < \const{1000}; i++) \{} \\
		\texttt{~~~~~~~~\flow{if} ((status = \call{\_xbegin}()) == \const{\_XBEGIN\_STARTED}) \{} \\
		\texttt{~~~~~~~~~~~~x++;} \\
		\texttt{~~~~~~~~~~~~\call{\_xend}();} \\
		\texttt{~~~~~~~~\} \flow{else} \{} \\
		\texttt{~~~~~~~~~~~~\call{mutex\_lock}(\&m);} \\
		\texttt{~~~~~~~~~~~~x++;} \\
		\texttt{~~~~~~~~~~~~\call{mutex\_unlock}(\&m);} \\
		\texttt{~~~~~~~~\}} \\
		\texttt{~~~~\}} \\
		\texttt{\}} \\
		\end{tabular}
	\end{center}
	\caption{The example {\tt count} routine from Figure~\ref{fig:mutex}, rewritten to use HTM.
		If the transaction in the top branch aborts,
		whether from a memory conflict or random system interrupt,
		%from the programmer's intention,
		execution will revert to the return of {\tt \_xbegin},
		{\tt status} will be assigned an error code indicating the abort reason,
		and control will drop into the {\tt else} branch.
		The programmer can then use explicit synchronization, such as a mutex, to resolve the conflict.}
	\label{fig:htm-example}
\end{figure}

@@

\begin{figure}[t]
        \small
	\begin{center}
\begin{tabular}{c}
\begin{tabular}{rll}
        & \multicolumn{2}{c}{\texttt{int x = 0; bool y = false; mutex\_t mx;}} \\
        \\
        & {\bf Thread 1} & {\bf Thread 2} \\
        1 & \texttt{\hilight{brickred}{x++;}~// A1} & \\
        2 & \texttt{mutex\_lock(\&mx);} & \\
        3 & \texttt{mutex\_unlock(\&mx);} & \\
        4 & & \texttt{mutex\_lock(\&mx);} \\
        5 & & \texttt{mutex\_unlock(\&mx);} \\
        6 & & \texttt{\hilight{brickred}{x++;}~// A2} \\
\end{tabular}
\\
\\
	{\normalsize (a) True potential data race.}
\\
\\
\begin{tabular}{rll}
        %& \multicolumn{2}{c}{\texttt{int x = 0; bool y = false; mutex\_t mx;}} \\
        & {\bf Thread 1} & {\bf Thread 2} \\
        1 & \texttt{\hilight{brickred}{x++;}~// B1} & \\
        2 & \texttt{mutex\_lock(\&mx);} & \\
        3 & \texttt{y = true;} & \\
        4 & \texttt{mutex\_unlock(\&mx);} & \\
        5 & & \texttt{mutex\_lock(\&mx);} \\
        6 & & \texttt{bool tmp = y;} \\
        7 & & \texttt{mutex\_unlock(\&mx);} \\
        8 & & \texttt{if (tmp) \hilight{brickred}{x++;}~// B2} \\
        %8 & & \texttt{if (tmp)} \\
        %9 & & \texttt{~~~~\hilight{brickred}{x++;}~// B2} \\
\end{tabular}
\\
\\
{\normalsize (b) No data race in any interleaving.}
\end{tabular}
	\end{center}
\caption{{Data-race analyses may be prone to either {\em false negatives} or {\em false positives}.
Applying Happens-Before to program (a) will miss the potential race possible between A1/A2 in an alternate interleaving,
while using Limited Happens-Before on (b) will produce a false alarm on B1/B2.}}
\label{fig:hb-example}
\end{figure}

>>>>

%\inspirationalquote{We pray to our ancestors. We've met all of them. None of them were better than anyone else.}
%{Mina, Iconoclasts}

\begin{flushright}
	{
	\fontspec[Path=./fonts/]{AlexaStd}
	\begin{tabular}{p{0.49\textwidth}}
	{\em "Well I don't *feel* brave. I feel scared, and insignificant."}
	\\
	{\em "That's what heroes always say, at first."}
	\end{tabular}
	}
	\vspace{2em}
\end{flushright}

@@

\begin{figure}[t]
	\begin{tabular}{cc}
		\begin{tabular}{p{0.45\textwidth}p{0.5\textwidth}}
			{\small
			\begin{tabular}{l}
				\texttt{\ctype{int} x;} \\
				\texttt{\ctype{void} \call{count}() \{} \\
				\texttt{~~~~\call{for} (\ctype{int} i = \const{0}; i < \const{1000}; i++)} \\
				\texttt{~~~~~~~~x++;} \\
				\texttt{\}} \\
				\texttt{\ctype{void} \call{main}() \{} \\
				\texttt{~~~~tid1 = \call{thr\_create}(count);} \\
				\texttt{~~~~tid2 = \call{thr\_create}(count);} \\
				\texttt{~~~~\call{thr\_join}(tid1);} \\
				\texttt{~~~~\call{thr\_join}(tid2);} \\
				\texttt{~~~~\call{printf}(\const{"\%d\textbackslash{}n"}, x);} \\
				\texttt{\}} \\
			\end{tabular}
			}
			&
			{\small \begin{tabular}{ll}
				{\bf Thread 1} & {\bf Thread 2} \\
				\hline
				\texttt{load tmp <- x;} & \\
				& \texttt{load tmp <- x;} \\
				& \texttt{add tmp <- 1;} \\
				& \texttt{store x <- tmp;} \\
				\texttt{add tmp <- 1;} & \\
				\texttt{store x <- tmp;} & \\
			\end{tabular}
			}
			\\
			\\
			(a) Source listing for a multithreaded program which might count to 2000.
			&
			(b) Example interleaving of the compiled assembly for (a),
			in which 2 concurrent iterations of the loop yield 1 net increment of {\tt x}.
		\end{tabular}
	\end{tabular}
	\caption{Example concurrent program in which simultaneous accesses to shared state may interleave to produce unexpected results.}
	\label{fig:concurrency-bug}
\end{figure}

@@

\begin{figure}[t]
	\begin{tabular}{p{0.45\textwidth}p{0.5\textwidth}}
		{\small
		\begin{tabular}{l}
			\texttt{\ctype{typedef struct mutex} \{} \\
			\texttt{~~~~\ctype{volatile int} held;} \\
			\texttt{~~~~\ctype{int} owner;} \\
			\texttt{\} \ctype{mutex\_t};} \\
			\texttt{\ctype{void} \call{mutex\_lock}(\ctype{mutex\_t} *mp) \{}\\
			\texttt{~~~~\flow{while} (\call{xchg}(mp->held, \const{1}))} \\
			\texttt{~~~~~~~~\call{yield}(mp->owner);} \\
			\texttt{~~~~mp->owner = \call{gettid}();} \\
			\texttt{\}} \\
			\texttt{\ctype{void} \call{mutex\_unlock}(\ctype{mutex\_t} *mp) \{}\\
			\texttt{~~~~mp->owner = \ctype{-1};} \\
			\texttt{~~~~mp->held = \ctype{0};} \\
			\texttt{\}} \\
		\end{tabular}
		}
		&
		{\small
		\begin{tabular}{l}
			\texttt{\ctype{int} x;} \\
			\texttt{\ctype{mutex\_t} m;} \\
			\texttt{\ctype{void} \call{count}() \{} \\
			\texttt{~~~~\flow{for} (\ctype{int} i = \const{0}; i < \const{1000}; i++) \{} \\
			\texttt{~~~~~~~~\call{mutex\_lock}(\&m);} \\
			\texttt{~~~~~~~~x++;} \\
			\texttt{~~~~~~~~\call{mutex\_unlock}(\&m);} \\
			\texttt{~~~~\}} \\
			\texttt{\}} \\
		\end{tabular}
		}
		\\
		(a) A simple mutual exclusion lock built using the {\tt xchg} instruction. %(accessed using a GCC compiler intrinsic).
		&
		(b) The {\tt count} function from Figure~\ref{fig:concurrency-bug}, adjusted to use a mutex to ensure each increment of {\tt x} is uninterruptible.
	\end{tabular}
	\caption{Using a locking primitive to protect accesses to shared state.}
	\label{fig:mutex}
\end{figure}

@@

\begin{figure}[t]
	\begin{center}
		\small
		\begin{tabular}{l}
		\texttt{\ctype{void} \call{count}() \{} \\
		\texttt{~~~~\flow{for} (\ctype{int} i = \const{0}; i < \const{1000}; i++) \{} \\
		\texttt{~~~~~~~~\flow{if} ((status = \call{\_xbegin}()) == \const{\_XBEGIN\_STARTED}) \{} \\
		\texttt{~~~~~~~~~~~~x++;} \\
		\texttt{~~~~~~~~~~~~\call{\_xend}();} \\
		\texttt{~~~~~~~~\} \flow{else} \{} \\
		\texttt{~~~~~~~~~~~~\call{mutex\_lock}(\&m);} \\
		\texttt{~~~~~~~~~~~~x++;} \\
		\texttt{~~~~~~~~~~~~\call{mutex\_unlock}(\&m);} \\
		\texttt{~~~~~~~~\}} \\
		\texttt{~~~~\}} \\
		\texttt{\}} \\
		\end{tabular}
	\end{center}
	\caption{The example {\tt count} routine from Figure~\ref{fig:mutex}, rewritten to use HTM.
		If the transaction in the top branch aborts,
		whether from a memory conflict or random system interrupt,
		%from the programmer's intention,
		execution will revert to the return of {\tt \_xbegin},
		{\tt status} will be assigned an error code indicating the abort reason,
		and control will drop into the {\tt else} branch.
		The programmer can then use explicit synchronization, such as a mutex, to resolve the conflict.}
	\label{fig:htm-example}
\end{figure}

@@

\begin{figure}[t]
        \small
	\begin{center}
\begin{tabular}{c}
\begin{tabular}{rll}
	& \multicolumn{2}{c}{\texttt{\ctype{int} x = \const{0}; \ctype{bool} y = \const{false}; \ctype{mutex\_t} mx;}} \\
        \\
        & {\bf Thread 1} & {\bf Thread 2} \\
	1 & \texttt{\hilight{brickred}{x++;}~\ccomment{// A1}} & \\
	2 & \texttt{\call{mutex\_lock}(\&mx);} & \\
	3 & \texttt{\call{mutex\_unlock}(\&mx);} & \\
	4 & & \texttt{\call{mutex\_lock}(\&mx);} \\
	5 & & \texttt{\call{mutex\_unlock}(\&mx);} \\
	6 & & \texttt{\hilight{brickred}{x++;}~\ccomment{// A2}} \\
\end{tabular}
\\
\\
	{\normalsize (a) True potential data race.}
\\
\\
\begin{tabular}{rll}
        %& \multicolumn{2}{c}{\texttt{int x = 0; bool y = false; mutex\_t mx;}} \\
        & {\bf Thread 1} & {\bf Thread 2} \\
	1 & \texttt{\hilight{brickred}{x++;}~\ccomment{// B1}} & \\
	2 & \texttt{\call{mutex\_lock}(\&mx);} & \\
	3 & \texttt{y = \const{true};} & \\
	4 & \texttt{\call{mutex\_unlock}(\&mx);} & \\
	5 & & \texttt{\call{mutex\_lock}(\&mx);} \\
	6 & & \texttt{\ctype{bool} tmp = y;} \\
	7 & & \texttt{\call{mutex\_unlock}(\&mx);} \\
	8 & & \texttt{\flow{if} (tmp) \hilight{brickred}{x++;}~\ccomment{// B2}} \\
        %8 & & \texttt{if (tmp)} \\
        %9 & & \texttt{~~~~\hilight{brickred}{x++;}~// B2} \\
\end{tabular}
\\
\\
{\normalsize (b) No data race in any interleaving.}
\end{tabular}
	\end{center}
\caption{{Data-race analyses may be prone to either {\em false negatives} or {\em false positives}.
Applying Happens-Before to program (a) will miss the potential race possible between A1/A2 in an alternate interleaving,
while using Limited Happens-Before on (b) will produce a false alarm on B1/B2.}}
\label{fig:hb-example}
\end{figure}
